---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rsniffs
<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-blue.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![CRAN status](https://www.r-pkg.org/badges/version/rsniffs)](https://CRAN.R-project.org/package=rsniffs)
[![](http://cranlogs.r-pkg.org/badges/grand-total/rsniffs?color=green)](https://cran.r-project.org/package=rsniffs)
[![](http://cranlogs.r-pkg.org/badges/last-week/rsniffs?color=green)](https://cran.r-project.org/package=rsniffs)
[![Codecov test coverage](https://codecov.io/gh/PaulESantos/rsniffs/branch/main/graph/badge.svg)](https://app.codecov.io/gh/PaulESantos/rsniffs?branch=main)
[![R-CMD-check](https://github.com/PaulESantos/rsniffs/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/PaulESantos/rsniffs/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->


## Overview

**`rsniffs`** is an R package that provides programmatic access to selected statistical datasets from the [National Forest and Wildlife Information System (SNIFFS)](https://sniffs.serfor.gob.pe/estadistica/es), managed by SERFOR (Servicio Nacional Forestal y de Fauna Silvestre) in Peru.

SNIFFS is a national platform designed to support decision-making in the forest and wildlife sector. Among its various components, this package focuses **exclusively** on the **Statistical Component**, providing structured and cleaned access to publicly available datasets.

> 🔗 For more information about SNIFFS, visit: [sniffs.serfor.gob.pe](https://sniffs.serfor.gob.pe)

---

## Features

This package allows users to access cleaned and up-to-date data from the following developed elements within the Statistical Component of SNIFFS:

| COMPONENT                                      | Dataset Element                                                                 | Status                          |
|------------------------------------------------|---------------------------------------------------------------------------|---------------------------------|
| TÍTULOS HABILITANTES                           | CONCESIONES                                                               | Desarrollada                    |
| TÍTULOS HABILITANTES                           | AUTORIZACIONES                                                            | Desarrollada como data set      |
| TÍTULOS HABILITANTES                           | PERMISOS                                                                  |                                 |
| TÍTULOS HABILITANTES                           | BOSQUES LOCALES                                                           | Desarrollada                    |
| TÍTULOS HABILITANTES                           | CESIÓN EN USO EN SISTEMA AGROFORESTAL                                     |                                 |
| REGISTROS NACIONALES                           | REGENTES FORESTALES Y DE FAUNA SILVESTRE                                 |                                 |
| REGISTROS NACIONALES                           | PLANTACIONES FORESTALES                                                  |                                 |
| REGISTROS NACIONALES                           | ESPECIALISTAS                                                             |                                 |
| REGISTROS NACIONALES                           | REGISTRO NACIONAL DE INFRACTORES                                         |                                 |
| FAUNA SILVESTRE                                | CURSO DE EDUCACIÓN, SEGURIDAD Y ÉTICA EN LA CAZA DEPORTIVA               | Desarrollada                    |
| FAUNA SILVESTRE                                | REGISTRO DE EMPRESAS AUTORIZADAS                                         | Desarrollada                    |
| FAUNA SILVESTRE                                | CENTRO DE MANEJO/CRÍA DE FAUNA SILVESTRE                                 |                                 |
| AUTORIZACIONES                                 | DESBOSQUE                                                                 |                                 |
| AUTORIZACIONES                                 | CAMBIO DE USO                                                             |                                 |
| PRODUCCIÓN, INDUSTRIA Y COMERCIO               | PRODUCCIÓN FORESTAL                                                      |                                 |
| PRODUCCIÓN, INDUSTRIA Y COMERCIO               | DEPÓSITOS Y ESTABLECIMIENTOS                                             | Desarrollada                    |
| PRODUCCIÓN, INDUSTRIA Y COMERCIO               | CENTROS DE TRANSFORMACIÓN                                                |                                 |
| PRODUCCIÓN, INDUSTRIA Y COMERCIO               | PRECIOS FORESTALES                                                       |                                 |
| PRODUCCIÓN, INDUSTRIA Y COMERCIO               | BAMBÚ A NIVEL NACIONAL                                                   | Desarrollada                    |
| CAMÉLIDOS SUDAMERICANOS SILVESTRES             | CAMÉLIDOS SUDAMERICANOS                                                  |                                 |
| CAMÉLIDOS SUDAMERICANOS SILVESTRES             | DECLARACIONES DE MANEJO                                                  |                                 |
| COMERCIO EXTERIOR                              | EXPORTACIONES                                                             |                                 |
| COMERCIO EXTERIOR                              | PERMISOS DE EXPORTACIÓN                                                  | Desarrollada                    |
| CONTROL DE PRODUCTORES                         | PRODUCTOS TRANSFERIDOS                                                   |                                 |
| CONTROL DE PRODUCTORES                         | SEDES Y PUESTOS DE CONTROL                                               | Desarrollada                    |
| CONTROL DE PRODUCTORES                         | PRODUCTORES FORESTALES CLASIFICADOS PARA SER TRANSFERIDOS               |                                 |
| INVESTIGACIÓN CIENTÍFICA Y RECURSOS GENÉTICOS  | AUTORIZACIONES DE INVESTIGACIÓN CIENTÍFICA                               |                                 |
| INVESTIGACIÓN CIENTÍFICA Y RECURSOS GENÉTICOS  | REGISTRO DE ACCESO A RECURSOS GENÉTICOS                                  |                                 |
| INVESTIGACIÓN CIENTÍFICA Y RECURSOS GENÉTICOS  | INSTITUCIONES CIENTÍFICAS NACIONALES DEPOSITARIAS DE MATERIAL BIOLÓGICO | Desarrollada                    |


Each function downloads the data directly from official SERFOR Google Drive spreadsheets, processes and cleans the information.

---

## Installation

You can install the development version of `rsniffs` from GitHub using [`pak`](https://pak.r-lib.org/) or [`remotes`](https://github.com/r-lib/remotes):

```r
# Using pak
pak::pak("PaulESantos/rsniffs")

# Or using remotes
remotes::install_github("PaulESantos/rsniffs")
```


## Usage

```{r}
library(rsniffs)

# Example: Get bamboo  data
pic_plantaciones_bambu()
pic_movilizacion_bambu()
```

## Available Functions

| Function                      | Description                                                              |
|------------------------------|--------------------------------------------------------------------------|
| `th_concesiones()`           | Get registry of forest concessions                                       |
| `th_autorizaciones`          | Internal dataset on authorizations                                       |
| `th_bosques_locales()`       | Get registry of local forests                                            |
| `fs_curso_caza_deportiva()`  | Get registry of authorized hunting courses                               |
| `fs_empresas_autorizadas()`  | Get registry of authorized wildlife companies                            |
| `pi_depositos()`             | Get registry of deposits and establishments                              |
| `pi_bambu_nacional()`        | Get bamboo-related data at the national level                            |
| `ce_permisos_exportacion()`  | Get registry of issued export permits                                    |
| `cp_sede_puestos_control()`  | Get list of forest and wildlife control offices                          |
| `icrg_inst_depositarias()`   | Get list of national scientific institutions storing biological samples  |


## Data Sources

 All data is sourced from official SERFOR datasets published through the SNIFFS statistical portal. These datasets are public, periodically updated, and hosted on Google Drive.

## Notes
📶 An active internet connection is required to retrieve the data.

⚠️ All datasets are read from Google Sheets hosted by SERFOR and may change structure without prior notice.

## Contributing
If you'd like to contribute or suggest improvements, feel free to open an issue or submit a pull request on GitHub.

## Citation
If you use rsniffs in your work, please cite it as follows:
```r
Santos Andrade, Paul E. (2025). rsniffs: Access Statistical Data from the National Forest and Wildlife Information System (SNIFFS), Peru.
```
